### 2022-07-08 chapters/optimization/python-dir/python.tex
add : ml figures

When it comes to developing software systems, one can notice that---except Perl---dynamic programming languages have taken over compiled languages along the last decade in terms of developers popularity (cf. Figure~\ref{fig:pypl}).
However, it remains unclear if this class of dynamic programming languages can reasonably compete with compiled ones when it comes to power efficiency.

\begin{figure}[htbp]
    \includegraphics[width=\linewidth]{imgs/programminglanguangespopularity.png}
    \caption{PYPL Popularity of Programming Languages~\cite{noauthor_pypl_nodate}}
    \label{fig:pypl}
\end{figure}

In particular, Noureddine~\emph{et~al.}~\cite{noureddine_preliminary_2012} in 2012, and then Pereira~\emph{et~al.}~\cite{pereira_energy_2017} in 2017, conducted empirical power measurements on this topic, and both concluded that compiled programming languages overcome dynamic ones when it comes to power consumption.
According to their experiments, an interpreted programming language like Python can impose up to a $7,588\,\%$ energy overhead compared to C~\cite{pereira_energy_2017} (cf. Figure~\ref{fig:hannoi}).
In this chapter, we therefore explore the oblivious optimizations that can be applied to Python legacy applications in order to reduce their energy footprint.
As Python is widely adopted by software services deployed in public and private Cloud infrastructures, we believe that our contributions will benefit to wide diversity of legacy systems and not only favorably contribute to reduce the carbon emissions of ICT, but also reduce their cloud invoice for the resources consumed by these services.
More specifically, this chapter focuses on runtime optimizations that can adopted by developers to leverage the power consumption of Python applications.
We start by studuing the impact of some of programmers choice such as the type of data structures or loops, on the global energy consumption of the execution code.
then we discuss some other factors such as level of concurrency etc. later we we will talk about another non intrusive ways to optimize the energy consumption of this code. one kind of those optimizations includes alternative interpertersand some libraries that are dedicated to optimize the code without changing its structures such as \emph{ahead-of-time} (AOT) compilation and \emph{just-in-time} (JIT) libraries that are maintained by the community.
% However, the operational conditions under which these measurements may threaten the validity of their results.

@@

\begin{figure}[thb]
    \centering
    \includegraphics[width=\linewidth]{imgs/python-implementations-tree}
    \caption{Python interpeters}
    \label{fig:interpreters}
\end{figure}

@@

%%% putting the idea, later i ll reformulate it 
the reason behind a such behaviour is mainly related on how python interpeter is made.
% add refenece https://www.python.org/doc/essays/list2str/ 
%% https://www.pythonpool.com/for-vs-while-loop-python/  
% https://wiki.python.org/moin/TimeComplexity
%% http://python-history.blogspot.com/2010/06/from-list-comprehensions-to-generator.html?m=1
In order to reduce the latency of the python code. most of the buitl-in functions an operations are written in C, same thing goes for the function range, further more the function len has a complexity of O(1) because it is based on the function Py\_SIZE  of C which stores the length in a field for the object.

so basically for the for in range is basically creating a new iterator that has the same length of the first one. and for each iteration we will have to do a second acceess ( l[i]) instead of one one access hence the double time.

for the wile is even slower due to the implicite incrementation of the variable which will cause an extra operation during the loop .

to confirm the hyphothsis. we tried to construct a new list by editing the elements of the previous one. And as predicted , the builtin methodes are the best energy saving, while the customize while loop is the heaviest. an interesting finding is. the impact of annonymous functions ( lambda expressions) on the energy consumption.
the reseason behind this is the fact that python treats those functions as local variables unlike the predifined ones which are global in our case. therefore they are faster and consumes less energy  % same reference https://www.python.org/doc/essays/list2str/

\begin{figure}[htb]
    \centering
    \includegraphics[width=\linewidth]{imgs/python_iterations}
    \caption{energy behaviour of different python loops }
    \label{fig:pythonloops}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=\linewidth]{imgs/python_treatemens}
    \caption{energy behaviour of different methodes to change a list}
    \label{fig:pythontreatement}
\end{figure}

@@

@@

@@

\begin{figure}[thb]
    \centering
    \includegraphics[width=\linewidth]{imgs/bitopts_mean}
    \caption{Mean consumption of different implementations of bit operations (Joule) }
    \label{fig:bitops}
\end{figure}

\begin{figure}[thb]
    \centering
    \includegraphics[width=\linewidth]{imgs/alltomti_performance}
    \caption{different interpererts optimisation }
    \label{fig:tommi_all}
\end{figure}
\begin{figure}[thb]
    \centering
    \includegraphics[width=\linewidth]{imgs/tommti_compare__pypy2_pypy3}
    \caption{green factor of pypy }
    \label{fig:pypy2vspypy3}
\end{figure}

\begin{figure}[thb]
    \centering
    \includegraphics[width=\linewidth]{imgs/tommti_compare__cpython3_numba3_pypy3}
    \caption{comparaison of pypy vs python vs numba }
    \label{fig:p3}
\end{figure}

\begin{figure}[thb]
    \centering
    \includegraphics[width=\linewidth]{imgs/tommti_compare__cpython2_pypy2_numba2}
    \caption{comparaison of pypy vs python vs numba }
    \label{fig:p2}
\end{figure}

@@

\section{conclusion}

one can notice that the choice of the right interpeter can impact havely the energy consumption of the python programs. The fact that there is no general solution make this study more interesting. however the main drawback is the lack of the compatibility for some of those solutions which lead to some sacrifices when we want something generic

>>>>

Dynamic programming languages, with the exception of Perl, have surpassed compiled programming languages in terms of popularity among software system developers over the past decade (cf. Figure~\ref{fig:pypl}).
However, it remains unclear whether this category of dynamic programming languages can truly compete with compiled ones in terms of power consumption. 
\begin{figure}[htbp]
    \includegraphics[width=\linewidth]{imgs/programminglanguangespopularity.png}
    \caption{PYPL Popularity of Programming Languages~\cite{noauthor_pypl_nodate}}
    \label{fig:pypl}
\end{figure}

In particular, Noureddine~\emph{et~al.}~\cite{noureddine_preliminary_2012} in 2012, and then Pereira~\emph{et~al.}~\cite{pereira_energy_2017} in 2017, conducted empirical power measurements on this topic, and both concluded that compiled programming languages overcome dynamic ones when it comes to power consumption.
According to their experiments, an interpreted programming language like Python can impose up to a $7,588\,\%$ energy overhead compared to C~\cite{pereira_energy_2017} (cf. Figure~\ref{fig:hannoi}).
In this chapter, we therefore explore the oblivious optimizations that can be applied to Python legacy applications in order to reduce their energy footprint.
As Python is widely adopted by software services deployed in public and private Cloud infrastructures, we believe that our contributions will benefit to wide diversity of legacy systems and not only favorably contribute to reduce the carbon emissions of ICT, but also reduce their cloud invoice for the resources consumed by these services.
More specifically, this chapter focuses on runtime optimizations that can adopted by developers to leverage the power consumption of Python applications.
We start by studuing the impact of some of programmers choice such as the type of data structures or loops, on the global energy consumption of the execution code.
then we discuss some other factors such as level of concurrency etc. later we we will talk about another non intrusive ways to optimize the energy consumption of this code. one kind of those optimizations includes alternative interpertersand some libraries that are dedicated to optimize the code without changing its structures such as \emph{ahead-of-time} (AOT) compilation and \emph{just-in-time} (JIT) libraries that are maintained by the community.

%%%%%%%%%%% READ AGAIN 
articular, Noureddine in 2012 and Pereira in 2017 conducted empirical power measurements on this topic, and both concluded that compiled programming languages consume less power than dynamic ones.They found that a programming language that is interpreted, like Python, can use up to 7,584 times more energy than C. As a result, in this chapter, we will look at the obvious optimizations that may be done to Python legacy programs to lower their energy footprint.Since Python is used by many software services in both public and private cloud infrastructures, we think that our contributions will help a wide range of legacy systems by reducing ICT carbon emissions and lowering cloud bills for the resources used by these services. This chapter focuses on runtime improvements that developers can employ to reduce the power consumption of Python applications.We start by looking at how the choices made by programmers, like the type of data structure or loop, affect how much energy the code as a whole uses. Then we discuss additional topics, such as the level of concurrency, and so on.We will discuss another non-intrusive technique to optimize this code's energy consumption.Alternative interpreters and libraries dedicated to optimizing code without affecting its structures, such as ahead-of-time (AOT) compilation and just-in-time (JIT) libraries maintained by the community, are examples of these optimizations. 
% However, the operational conditions under which these measurements may threaten the validity of their results.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/python-implementations-tree}
    \caption{Python interpeters}
    \label{fig:interpreters}
\end{figure}

@@

%%% putting the idea, later i ll reformulate it 
the reason behind a such behaviour is mainly related on how python interpeter is made.
% add refenece https://www.python.org/doc/essays/list2str/ 
%% https://www.pythonpool.com/for-vs-while-loop-python/  
% https://wiki.python.org/moin/TimeComplexity
%% http://python-history.blogspot.com/2010/06/from-list-comprehensions-to-generator.html?m=1
In order to reduce the latency of the python code. most of the buitl-in functions an operations are written in C, same thing goes for the function range, further more the function len has a complexity of O(1) because it is based on the function Py\_SIZE  of C which stores the length in a field for the object.
so basically for the for in range is basically creating a new iterator that has the same length of the first one. and for each iteration we will have to do a second acceess ( l[i]) instead of one one access hence the double time.
for the wile is even slower due to the implicite incrementation of the variable which will cause an extra operation during the loop .
to confirm the hyphothsis. we tried to construct a new list by editing the elements of the previous one. And as predicted , the builtin methodes are the best energy saving, while the customize while loop is the heaviest. an interesting finding is. the impact of annonymous functions ( lambda expressions) on the energy consumption.
the reseason behind this is the fact that python treats those functions as local variables unlike the predifined ones which are global in our case. therefore they are faster and consumes less energy  % same reference https://www.python.org/doc/essays/list2str/

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/python_iterations}
    \caption{energy behaviour of different python loops }
    \label{fig:pythonloops}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/python_treatemens}
    \caption{energy behaviour of different methodes to change a list}
    \label{fig:pythontreatement}
\end{figure}

@@

\subsection{Python and machine learning }
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/accuracy_basedonepoch}
    \caption{accuracy based on epoch  }
    \label{fig:p2}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/power_gpu_baedonepoche}
    \caption{evolution of average gpu power based on epoch  }
    \label{fig:p2}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/cumulative_energy_fast10}
    \caption{cumulative energy of fast10 benchmark within different configurations }
    \label{fig:p2}
\end{figure}

\paragraph{conclusion}
We discovered that when a model's accuracy improves, so does the energy required for a subsequent accuracy gain.
This begs the question of when we should discontinue training.
Is a 10\% increase in accuracy worth it if we have to spend three times the energy?
Even while utilizing a GPU consumes more total power, the training time is significantly reduced.
As a result, the utilization of a GPU is required to lower the energy consumption of the training.
Some experiments revealed unusual behavior that was not studied during this internship.
These could be examined in future studies and perhaps used to achieve more energy savings.
The execution time did not depend on the number of cores employed in several studies.
This parameter had a significant impact on training time in others.
As a result, the best core configuration will be determined by the script.
The next steps in this area could include exploring the reasons for these discrepancies and possibly creating a script that can automatically find the ideal core configuration for a specific script.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/bitopts_mean}
    \caption{Mean consumption of different implementations of bit operations (Joule) }
    \label{fig:bitops}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/alltomti_performance}
    \caption{different interpererts optimisation }
    \label{fig:tommi_all}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/tommti_compare__pypy2_pypy3}
    \caption{green factor of pypy }
    \label{fig:pypy2vspypy3}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/tommti_compare__cpython3_numba3_pypy3}
    \caption{comparaison of pypy vs python vs numba }
    \label{fig:p3}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/tommti_compare__cpython2_pypy2_numba2}
    \caption{comparaison of pypy vs python vs numba }
    \label{fig:p2}
\end{figure}

@@

@@

\section{conclusion}
One may observe that the choice of Python interpreter has a significant impact on the programs' energy consumption.
This investigation is made more intriguing by the absence of a universal solution.
The primary downside is the incompatibility of some of these solutions, which causes us to make concessions when we need a generic answer.
